---
title: "Lab 6"
author: "Ishaan Phadnis"
date: "November 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  4 (Pg. 302); 1 (Pgs. 316-317); 1 (Pgs. 328-329); 1, 2 (Pgs. 353-354)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 16 (Pg. 302)

##  Exercise 4

### a)
```{r}
#
function_last = function(x){
  x[[length(x)]]
}

function_last(3:7)
```

### b)
```{r}
#
function_even = function(x) {
  x[seq(2, length(x), by = 2)]
}

function_even(5:13)
```

### c)
```{r}
#
function_notlast = function(x) {
  x[-length(x)]
}

function_notlast(2:9)
```

### d)
```{r}
#
function_even2 = function(x) {
  x[!is.na(x) & x %% 2 == 0]
}

function_even2(sample(c(NA, 10:50), 40))
```

# Chapter 17 (Pgs. 316-317)

##  Exercise 1

### a)
```{r}
#
mean_mtcars = vector("double", ncol(mtcars))
names(mean_mtcars) = names(mtcars)

for (i in names(mtcars)) {
  mean_mtcars[i] = mean(mtcars[[i]])
}

mean_mtcars
```

### b)
```{r}
#
flight_type = vector("list", ncol(nycflights13::flights))
names(flight_type) = names(nycflights13::flights)

for (i in names(nycflights13::flights)) {
  flight_type[[i]] = class(nycflights13::flights[[i]])
}

flight_type
```

### c)
```{r}
#
data("iris")
iris_unique = vector("double", ncol(iris))
names(iris_unique) = names(iris)

for (i in names(iris)) {
  iris_unique[i] = length(unique(iris[[i]]))
}
iris_unique
```

### d)
```{r}
#
x = 10
mean1 =  c(-10, 0, 10, 100)
rand_norm = vector("list", length(mean1))

for (i in seq_along(rand_norm)) {
  rand_norm[[i]] = rnorm(x, mean = mean1[i])
  }

rand_norm
```

# Chapter 17 (Pgs. 328-329)

##  Exercise 1

### a)
```{r}
#
map_dbl(mtcars, mean)
```

### b)
```{r}
#
map_chr(nycflights13::flights, typeof)
```

### c)
```{r}
#
map_int(iris, function(x) length(unique(x)))
```

### d)
```{r}
#
map(c(-10, 0, 10, 100), ~rnorm(n = 10, mean = .))
```

# Chapter 18 (Pgs. 353-354)

##  Exercise 1
```{r}
#
sim1a <- tibble(x = rep(1:10, each = 3), y = x * 1.5 + 6 + rt(length(x), df = 2))

ggplot(sim1a, aes(x = x, y = y)) + geom_point() + geom_smooth(method = "lm", se = FALSE)

```

Based on the simulations ran through the code above, there are no significant outliers in any of the generated plots and their slopes are pretty similar as well. 

##  Exercise 2
```{r}
#
measure_distance <- function(mod, data) {
  diff <- data$y - make_prediction(mod, data)
  mean(abs(diff))
}

make_prediction = function(mod, data) {
  mod[1] + mod[2] * data$x
}

param = optim(c(0, 0), measure_distance, data = sim1a)
param$par

```
