---
title: "Lab 5"
author: "Ishaan Phadnis"
date: "October 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pg. 227); 2 (Pg. 232); 1,2 (Pg. 235); 

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 12 (Pg. 227)

##  Exercise 1
```{r}
#
gss_cat %>% ggplot(aes(rincome)) + geom_bar()
gss_cat %>% mutate(rincome = fct_relevel(rincome, c("No answer", "Don't know", "Refused", "Not applicable"))) %>% ggplot(aes(rincome)) + geom_bar() + coord_flip()
```

So the code for 'levels' should be changed so that all non-income categories are at the end. Also, the plot is set to 'coord_flip' and this allows the labels to be read.

##  Exercise 3
```{r}
#
gss_cat %>% count(relig, denom) %>% filter(denom == "No denomination")

gss_cat %>% filter(!denom %in% c("No denomination", "No answer", "Other", "Don't know", "Not applicable")) %>% count(relig)

gss_cat %>% count(relig, denom) %>% ggplot(aes(x = relig, y = denom, size = n)) + geom_point() 
```

The 'relig' that 'denom' applies to is Protestant. This can be found by filtering for the non-responses, no answers, others, not-applicable, or no denomination. Also, a scatter plot of 'relig' vs. 'denom' can also be created.

# Chapter 12 (Pg. 232)

##  Exercise 2
```{r}
#
keep(gss_cat, is.factor) %>% names()
levels(gss_cat[["marital"]])
levels(gss_cat[["rincome"]])
levels(gss_cat[["partyid"]])
levels(gss_cat[["relig"]])
levels(gss_cat[["denom"]])
```

- The order of 'marital' is principled
- The order of 'race' is principled 
- The order of 'rincome' is arbitrary
- If you ignore “No answer”, “Don’t know”, and “Other party”, the order for 'partyid' is principled
- The order of 'relig' is arbitrary
- The order of 'denom' is also arbitrary 

# Chapter 12 (Pg. 235)

##  Exercise 1
```{r}
#
gss_cat %>% mutate(partyid = fct_collapse(partyid, other = c("No answer", "Don't know", "Other party"), rep = c("Strong republican", "Not str republican"), ind = c("Ind,near rep", "Independent", "Ind,near dem"), dem = c("Not str democrat", "Strong democrat"))) %>% count(year, partyid) %>% group_by(year) %>% mutate(p = n / sum(n)) %>% ggplot(aes(x = year, y = p, colour = fct_reorder2(partyid, year, p))) + geom_point() + geom_line() + labs(colour = "Party ID.")
```

The plot above shows the year-to-year trends of the different parties. 

##  Exercise 2
```{r}
#
library("stringr")

gss_cat %>%mutate(rincome = fct_collapse(rincome, `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"), `Lt $5000` = c("Lt $1000", str_c("$", c("1000", "3000", "4000"), " to ", c("2999", "3999", "4999"))),`$5000 to 10000` = str_c("$", c("5000", "6000", "7000", "8000"), " to ", c("5999", "6999", "7999", "9999")))) %>% ggplot(aes(x = rincome)) + geom_bar() + coord_flip()
```

The plot above collapses 'rincome' into different categories.