---
title: "Lab 1"
author: "Ishaan Phadnis"
date: "August 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense? 

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```

There are 234 rows and 11 columns in the dataset **mpg**. 

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Exercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Exercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
```{r}
ggplot(data = mpg) +
 geom_point(
 mapping = aes(x = displ, y = hwy, color = "blue"))
#
```

The reason why the points are not the color blue is because there should be a parenthesis between 'y = hwy' and the comma that follows. 

##  Exercise 2
```{r}
?mpg 
#
```

Variables that are categorical are Model, cyl, Manufacturer, trans, drv, fl, and class. Variables that are continuous are displ, year, cty, and hwy. Information is found in the help tab at the bottom right of RStudio. 

##  Exercise 5
```{r}
?geom_point
#
```

Stroke is an aesthetic that controls the width of the border of certain shapes and it can only be used with shapes that have a border. 

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
The line chart uses geom_line(), the box plot uses geom_boxplot(), the histogram uses geom_histogram(), and the area chart uses geom_area(). 

##  Exercise 2
```{r}
ggplot(
 data = mpg,
 mapping = aes(x = displ, y = hwy, color = drv)
) +
 geom_point() +
 geom_smooth(se = FALSE)
#
```

My head: The code looks like it will create some sort of plot, probability with several plotted points. 
Actuality: The code does create a graph of some sort that displays three different variables being displayed, all marked with a different color. Each variable is plotted with several different points and a regression line shows the trend of these points. 

##  Exercise 3
```{r}
show.legend = FALSE
```

This removes the legend. When removed, it essentially creates a more organized plot when the grouping is on a specific variable. 

##  Exercise 4
The se argument removes the intervals from the lines on the plot. 

##  Exercise 5
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised. The reason why is because the functions may look different by syntactically, but they both carry out the same exact processes. 

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
ggplot(data=diamonds) + geom_boxplot(aes(x= price, y = carat, color = color)) + facet_wrap(~clarity) + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + guides(color = guide_legend(title = "Color")) + xlab("") + ylab("Price") 

```
  
  


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 

```{r}
ggplot(data=diamonds) + geom_point(aes(x = carat, y = price, color = color, shape = clarity)) 
```

